---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rx86

<!-- badges: start -->
<!-- badges: end -->

The goal of rx86 is to provide a functional representation of x86 assembly 
and execute the instructions.

## Installation

You can install the development version of rx86 from Github with:

``` r
# install.packages("remotes")
remotes::install_github("jonocarroll/rx86")
```

## Motivation

See [this](https://jcarroll.com.au/2021/12/23/adventures-in-x86-asm/) blogpost, but I wanted to be able to 'run' some x86 assembly code and 
wondered if I could do it entirely from R.

A small, more general example, is to perform multiplication
```{r example}
library(rx86)

## load some asm instructions which store the 
## the numbers 167 and 28, multiplies them, 
## and stores the result in 0x22
mult_asm <- suppressWarnings(
  readr::read_fwf(system.file("asm", "mult.asm", package = "rx86"), 
                  col_types = "ccc",
                  col_positions = readr::fwf_widths(c(3, 6, 20)))
)
colnames(mult_asm) <- c("addr", "bytecode", "instr")
print(mult_asm)

# create blank slate of memory and registers
mem <- create_mem(len = 64)
registers <- create_reg()

runasm(mult_asm)

# extract result from memory and convert to integer
as.integer(mem[sanitize(0x22)])

167*28
```

This is detailed in `vignette("mult_code_petzold", package = "rx86")`.

What test is complete without a helloworld?
```{r}
hello_asm <- suppressWarnings(
  readr::read_fwf(system.file("asm", "helloworld.asm", package = "rx86"), 
                  col_types = "ccc",
                  col_positions = readr::fwf_widths(c(3, 3, 20)))
)
colnames(hello_asm) <- c("addr", "bytecode", "instr")
print(hello_asm)

mem <- create_mem()
registers <- create_reg()

runasm(hello_asm)
```

This is detailed in `vignette("helloworld", package = "rx86")`.

## Inspiration

After reading ['Code: The Hidden Language of Computer Hardware and Software'](https://en.wikipedia.org/wiki/Code:_The_Hidden_Language_of_Computer_Hardware_and_Software) by Charles Petzold - which does a fantastic job of describing 
the journey from telegraph relays through to a working computer - I was reminded that 
I had once looked into some assembly for a puzzle, but never wrote up my findings.

In an effort to better understand the solution to that puzzle (and the way to get to it) I thought it would be neat to try to replicate some of the functionality in a language I'm more familiar with. The operations on data stored in memory are simple - move a value from here to there, add some values, ... so it made sense to me to 
see if I could replicate that in a functional language.

This isn't something you would use to write asm, but it very well may help you understand what it's doing.

## Limitations

The following is a non-exhaustive list of the known limitations of this package.

* This was originally built to support a single use-case of solving the puzzle described in 
``vignette("dsd_ruxcon_challenge", package = "rx86")` so only has support for the
opcodes required for that. That covers a lot of possibilities, but if you need a 
`DIV` you're out of luck so far.
* The operations supported do so _in spirit_ and are not exact reproductions of 
functionality - if this does something different to real asm, that's not too surprising.
* This takes as input a `data.frame` of disassembled code. This has only been tested
with the output of 
```{bash, eval = FALSE}
udcli -x file.hex
```
with the `udis` tool from http://udis86.sourceforge.net/.
* 

## Other R assemblers
 * [c64](https://github.com/coolbutuseless/r64) - a c64/6502 assembler
