---
title: "Australian Defence Signals Directorate Ruxcon 2011 Challenge"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Australian Defence Signals Directorate Ruxcon 2011 Challenge}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rx86)
```

At ruxcon 2011, the following code was shown on a banner for the Australian (then named) Defence Signals Directorate
```{r}
dsd <- "6AAAAABbi8uDwx4zwDPSigOK
ETLCiAM8AHQrg8EBg8MB6+wz
/7/z+TEct0SlpGf5dRyl53US
YQEE56Ri7Kdkj8IAABkcOsw="
```
This is base-64 encoded, so let's decode it

```{r}
(b64 <- base64enc::base64decode(dsd))
```

E8 is an x86 CALL instruction - if we assume this is bytecode and disassemble it (e.g. with udis, which needs to be [installed](http://udis86.sourceforge.net/))
```{r}
(disas <- system("udcli -x", input = paste(b64, collapse = " "), intern = TRUE))
```

then read this as a data.frame
```{r}
asm <- suppressWarnings(
  readr::read_fwf(paste(disas, collapse = "\n"), 
                  col_types = "ccc",
                  col_positions = readr::fwf_widths(c(16, 16, 21)))
)
colnames(asm) <- c("addr", "bytecode", "instr")
# trim the leading 0s from addr since this is all we're using
asm$addr <- substr(asm$addr, nchar(asm$addr)-1, nchar(asm$addr))
asm
```
This appears to be a sequential operation from `0x00` to `0x21` with a loop back to 
`0xf` until `al == 0x0` at which point it jumps to the end at `0x46`. 
`int3` is an interrupt but let's generalise it to a `halt` because we'll be done
```{r}
asm[33, "instr"] <- "halt"
```
We can run this with `rx86`... we need a memory array and some registers
```{r}
mem <- create_mem()
registers <- create_reg()
```
Then we can run the code
```{r}
runasm(asm)
```
The 'code' part of the asm is stored in `0x00` to `0x21` with the remaining addresses being used for temporary storage, from `0x23`. The operations encoded 
perform an `XOR` between the values stored at `0x05` through to `0x21` with those starting at `0x23`, storing the results starting at `0x23`.
Extracting the memory from this offset onwards (up to where it zeroes) results in
```{r}
(mem_offset <- mem[which(names(mem) == "0x23"):length(mem)])
hex2string(mem_offset)
```
