---
title: "Hello, World!"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hello, World!}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rx86)
```

Of course, we need a helloworld example. Most of the asm examples use the 
human-readable version with a string and system calls - that's too fancy for me. 

Here I have hand-crafted the bytecode which outputs (via `int 0x80` - a custom
function which prints whatever is in the register `eax` to screen as character) 
the appropriate characters which are read directly from memory.

I wrote each line of this asm myself, so I really feel like I've actually "programmed" a computer for once, with some odd combination of zero dependencies (it's machine code!) and waaay too many (I use an R package to evaluate it).

The code is stored in this package. It stores the raw hex values for the characters of "Hello, world!" in the memory addresses starting at `0x08`. The 'code' is in the addresses prior to that, where the hex value for 14 (`0x0e`) is stored in `ecx`. `0x08` - the location of the first byte of the string, is moved to the accumulator `al`. The value at that address (the actual first byte of the string) is moved to `eax` which is the register from which the "system" will write to screen. The interrupt `int 0x80` signals this, and the first byte is converted to character and printed. `0x01` is subtracted from `ecx` (this is now counting how many characters remain to be printed) and if that results in zero (via the zero flag) then the code will jump to `0x17` which is a `halt`. Otherwise, `0x01` is added to the accumulator (holding the address of the next byte to read) and the code jumps back to `0x02`, loading the next byte into `eax`.
```{r}
hello_asm <- suppressWarnings(
  readr::read_fwf(system.file("asm", "helloworld.asm", package = "rx86"), 
                  col_types = "ccc",
                  col_positions = readr::fwf_widths(c(3, 3, 20)))
)
colnames(hello_asm) <- c("addr", "bytecode", "instr")
print(hello_asm)
```
As always, set up memory and registers
```{r}
mem <- create_mem()
registers <- create_reg()
```
Then we can run the code - this time we don't need to extract the memory
afterwards, the interrupt signals that the value in the `eax` register should be 
output to screen
```{r}
runasm(hello_asm)
```
