---
title: "Multiplying two integers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multiplying two integers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rx86)
```

Perform multiplication of 167 * 28 without a MUL opcode. This was inspired by 
['Code: The Hidden Language of Computer Hardware and Software'](https://en.wikipedia.org/wiki/Code:_The_Hidden_Language_of_Computer_Hardware_and_Software) by Charles Petzold where a 8-bit values were multiplied, requiring add-with-carry operations. For simplicity (and because I don't know how to emulate `hex(-1) == 0xff` without an actual 8-bit computer) I've converted it to just multiplying the two values without needing any carry registers. I have not updated the bytecode, so that is mostly meaningless.

Stores the value 167 in an accumulator as `0xa7` and adds `0xa7` to it 27 times
by storing a counter as `0x1c` and removing `0x01` from it each loop
until the zero flag is set. The result is stored in `0x22`.

The asm code for this is stored along with the package.
```{r}
mult_asm <- suppressWarnings(
  readr::read_fwf(system.file("asm", "mult.asm", package = "rx86"), 
                  col_types = "ccc",
                  col_positions = readr::fwf_widths(c(3, 6, 20)))
)
colnames(mult_asm) <- c("addr", "bytecode", "instr")
print(mult_asm)
```

In order to evaluate this code we require a memory array and some registers
```{r}
mem <- create_mem(len = 64)
registers <- create_reg()
```

The code can be evaluated, which will modify the memory array and registers
according to the operations
```{r}
runasm(mult_asm)
```

The final result can be extracted but it is still a hex value
```{r}
mem[sanitize(0x22)]
```

Converting it to an integer gives the expected result
```{r}
as.integer(mem[sanitize(0x22)])

167*28
```
